---
title: "Appalachian Gauges"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Site Selection:

```{r}
library(tidyverse)
library(ggforce)
library(dataRetrieval)
library(sf)
library(tmap)
library(spData)
library(repr)
library(zoo)
library(lubridate)
library(moments)
library(ContDataQC)
library(broom)
library(AICcmodavg)
library(raster)

#note: ecohydology doesn't play nice with other packages
#library(EcoHydRology)

tmap_mode("plot")

theme_set(theme_classic())

```

## Read in shapefile for Appalachian physiographic region

Excluding St. Lawrence Valley, New England Uplands, and Seaboard
Lowlands sections.

## Read in shapefile for gages

Taken from GAGES-II, intersected with physiographic region in ArcGIS.

```{r}
AppGages <- st_read("AppalachianGages_Filtered.shp")
  
AppRegions <- st_read("AppalachianRegions.shp")
```

## Filter to HCDN-2009 gages.

#### HCDN-2009 criteria:

(a) have 20 years of complete and continuous flow record in the last 20
    years (water years 1990-2009), and were thus also currently active
    as of 2009

(b) are identified as being in current reference condition according to
    the GAGES-II classification

(c) have less than 5 percent imperviousness as measured from the NLCD
    2006

(d) were not eliminated by a review from participating state Water
    Science Center evaluators.

```{r}
AppGages <- AppGages %>% filter(HCDN_2009 == "yes") %>% filter(PROVINCE != "NEW ENGLAND" & PROVINCE != "ADIRONDACK" & SECTION != "SOUTHERN NEW YORK" & SECTION != "CATSKILL")

```

## Map gages within physiographic provinces and sections.

```{r}
ProvinceMap <- tm_shape(AppRegions)+
  tm_polygons(col = "PROVINCE", palette = "Spectral")+
  tm_layout(legend.outside = TRUE)+
tm_shape(AppGages)+
  tm_dots()+
tm_shape(us_states)+
  tm_borders()

SectionMap <- tm_shape(AppRegions)+
  tm_polygons(col = "SECTION", textNA = "ADIRONDACK", palette = "Spectral")+
  tm_layout(legend.outside = TRUE)+
tm_shape(AppGages)+
  tm_dots()+
tm_shape(us_states)+
  tm_borders()

ProvinceMap
SectionMap
```

## Read in flow data availability from NWIS, join to gage table.

Note: whatNWISdata produces multiple rows per gage for each data type
and stat codes:

"dv" = daily values

"00003" = mean values

```{r}
sitenos <- c(AppGages$STAID)

availableData <- whatNWISdata(siteNumber = sitenos, parameterCd = "00060")

AppGages <- full_join(AppGages, availableData, by = c("STAID" = "site_no"))

AppGages <- AppGages %>% filter(data_type_cd == "dv", stat_cd == "00003")

```

## Create timeline of data availability

```{r, fig.height= 5}

timeline <- AppGages %>% ggplot(aes(x = begin_date, y = STAID, xend = end_date, yend = STAID))+
  geom_segment()
  #facet_wrap_paginate(facets = "PROVINCE", ncol = 1, nrow = 1, page = 1:5, scales = "free")

timeline
```

## Filter out and map gages excluded from period of interest

1981-2010: NWS normal period

```{r}
endLoss <- AppGages %>% filter(end_date <= "2010-12-31")

startLoss <- AppGages %>% filter(begin_date >= "1981-12-31")

SectionMap+
  tm_shape(startLoss)+
   tm_symbols(size = 0.2, col = "red")
  #tm_shape(endLoss)+
    #tm_symbols(size = 0.2, col = "red")

AppGages <- AppGages %>% filter(begin_date <= "1981-01-01" & end_date >= "2010-12-31")
```
