### This RMD contains the code to use the output of the Random Forest predictors to a linear multiple regression. 
# Using the top 10 important watershed characteristics for Flashiness, RBI, TenthP and NinetiethP from the random forest output

###Load libraries
```{r}
library(tidyverse)
# library(olsrr)
library(ggplot2)
library(Hmisc)
library(patchwork)
library(ggpmisc)
library(modelr)

```

### load in data
```{r}
forMR <- read_csv("Data/WSfactorsLarge.csv") #watershed characteristics

head(forMR)

RBItable <- read_csv("Data/FlashinessMean.csv") #pull Flashiness 

RBItable <- RBItable %>%
  dplyr::select(Flashiness, site_no) 

MRselect <- merge(forMR, RBItable, by.x = 'STAID', by.y = 'site_no')
```
###Correlation matrix
```{r}
# Remove top variables from RF for RBI
MRselect <- forMR %>%
  dplyr::select(BFI = BFI_AVE, HGC, slope_fdc, geol_porostiy, SANDAVE, HGB, RFACT, geol_permeability, soil_depth_statsgo, ROCKDEPAVE, aridity, ELEV_MEDIAN_M_BASIN, PLANTNLCD06, RIP100_DEV)

#correlation matrix
# variables with values above 0.7 (or highly correlated) discarded
cor_matrix <- cor(MRselect)

print(cor_matrix)

```



###Running BFI in a stepwise progression using top variables accross all top importance predictors
```{r}
# Check for missing and infinite values in the data
any(is.na(MRselect))

# Fit a stepwise regression model with the selected predictors
BFI_step_model <- step(lm(BFI ~ ., data = MRselect), direction = "both")

# Print the model summary
summary(BFI_step_model)

#copy best model lm call and use here to get coeffs if needed
# lm(formula = BFI ~ geol_porostiy + VarB + SANDAVE + HGB + soil_depth_statsgo + 
#     ROCKDEPAVE + runoff_ratio + ELEV_MEDIAN_M_BASIN, data = MRselect)

BFIpred <- MRselect %>%
  add_predictions(BFI_step_model)

BFI <- 
  BFIpred %>% drop_na() %>%
  ggplot(aes(x = BFI, y = pred)) +
  geom_point() +
  stat_poly_line(method=lm) +
  stat_poly_eq() +
  xlab("Actual BFI") +
  ylab("Predicted BFI") +
  theme_classic()

BFI

```
### RBI
```{r}
MRselect <- forMR %>%
  dplyr::select(RBI = Flashiness, HGC, slope_fdc, geol_porostiy, SANDAVE, HGB, RFACT, geol_permeability, soil_depth_statsgo, ROCKDEPAVE, aridity, ELEV_MEDIAN_M_BASIN, PLANTNLCD06, RIP100_DEV)

# Check for missing and infinite values in the data
any(is.na(MRselect))

# Fit a stepwise regression model with the selected predictors
RBI_step_model <- step(lm(RBI ~ ., data = MRselect), direction = "both")

# Print the model summary
summary(RBI_step_model)

#copy best model lm call and use here to get coeffs if needed
lm(formula = RBI ~ HGC + slope_fdc + geol_porostiy +  SANDAVE + 
    HGB + RFACT + geol_permeability + aridity + ELEV_MEDIAN_M_BASIN, 
    data = MRselect)

RBIpred <- MRselect %>%
  add_predictions(RBI_step_model)

RBI <- 
  RBIpred %>% drop_na() %>%
  ggplot(aes(x = RBI, y = pred)) +
  geom_point() +
  stat_poly_line(method=lm) +
  stat_poly_eq() +
  xlab("Actual Flashiness") +
  ylab("Predicted Flashiness") +
  theme_classic()

RBI
```

###Ninetieth Percentile flow
```{r}
NPselect <- forMR %>%
  dplyr::select(NP = NinetiethP, HGC, slope_fdc, geol_porostiy, SANDAVE, HGB, RFACT, geol_permeability, soil_depth_statsgo, ROCKDEPAVE, aridity, ELEV_MEDIAN_M_BASIN, PLANTNLCD06, RIP100_DEV) %>% drop_na()

# Check for missing and infinite values in the data
any(is.na(NPselect))

# Fit a stepwise regression model with the selected predictors
NP_step_model <- step(lm(NP ~ ., data = NPselect), direction = "both")

# Print the model summary
summary(NP_step_model)

#copy best model lm call and use here to get coeffs if needed
# lm(formula = NP ~ slope_fdc + VarB + SANDAVE + RFACT + soil_depth_statsgo + 
#     sand_frac + runoff_ratio + aridity + ELEV_MEDIAN_M_BASIN + 
#     RIP100_DEV, data = MRselect)

NPpred <- NPselect %>%
  add_predictions(NP_step_model)

NP <- 
  NPpred %>% drop_na() %>%
  ggplot(aes(x = NP, y = pred)) +
  geom_point() +
  stat_poly_line(method=lm) +
  stat_poly_eq() +
  xlab("Actual NinetiethP Flow (mm/day)") +
  ylab("Predicted NinetiethP Flow (mm/day)") +
  theme_classic()

NP
```

###Tenth Percentile flow
```{r}
TPselect <- forMR %>%
  dplyr::select(TP = TenthP, HGC, slope_fdc, geol_porostiy, SANDAVE, HGB, RFACT, geol_permeability, soil_depth_statsgo, ROCKDEPAVE, aridity, ELEV_MEDIAN_M_BASIN, PLANTNLCD06, RIP100_DEV)

# Check for missing and infinite values in the data
any(is.na(TPselect))

# Fit a stepwise regression model with the selected predictors
TP_step_model <- step(lm(TP ~ ., data = TPselect), direction = "both")

# Print the model summary
summary(TP_step_model)

#copy best model lm call and use here to get coeffs if needed
# lm(formula = TP ~ HGC + slope_fdc + geol_porostiy + VarB + HGB + 
#     RFACT + geol_permeability + soil_depth_statsgo + ROCKDEPAVE + 
#     runoff_ratio + aridity, data = TPselect)

TPpred <- TPselect %>%
  add_predictions(TP_step_model)

TP <- 
  TPpred %>% drop_na() %>%
  ggplot(aes(x = TP, y = pred)) +
  geom_point() +
  stat_poly_line(method=lm) +
  stat_poly_eq() +
  xlab("Actual TenthP Flow (mm/day)") +
  ylab("Predicted TenthP Flow (mm/day)") +
  theme_classic()

TP

MRplot <- BFI + RBI + NP + TP

MRplot

ggsave("MRplot.PNG", plot = MRplot)

```

