---
title: "Appalachian Gauges"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Site selection and data cleaning:
```{r, echo=FALSE}
library(tidyverse)
library(ggforce)
library(dataRetrieval)
library(sf)
library(tmap)
library(spData)
library(repr)
library(zoo)
library(lubridate)
library(moments)
library(broom)
library(AICcmodavg)
library(raster)
library(rgdal)

tmap_mode("plot")

theme_set(theme_classic())
```

## Read in shapefile for Appalachian physiographic region

Excluding St. Lawrence Valley, New England Uplands, and Seaboard
Lowlands sections.

## Read in shapefile for gages

Gages from CAMELS dataset, GAGESII shapefile joined to station by Gage ID

```{r}
AppRegions <- st_read("Data/Shapefiles/AppalachianRegions.shp")

AppGages <- read_csv("SAgages.csv")

AppGagesshp <- st_read("SAgages.shp") 

ProvinceMap <- tm_shape(AppRegions)+
  tm_polygons(col = "SECTION", palette = "Spectral")+
  tm_layout(legend.outside = TRUE)+
tm_shape(AppGagesshp)+
  tm_dots()+
tm_shape(us_states)+
  tm_borders()

ProvinceMap
head(AppGages)
```
## Join watershed attributes from CAMELS dataset:

 - Can add more watershed characteristics here.
```{r}
WSfactors <- full_join(read_delim("CAMELS Data/camels_clim.txt"), read_delim("CAMELS Data/camels_geol.txt"), by = "gauge_id") %>%
  full_join(read_delim("CAMELS Data/camels_soil.txt"), by = "gauge_id") %>%
  inner_join(AppGages, by = "gauge_id")

write_csv(WSfactors, "CAMELSapps_attr.csv")
```

## Testing CAMELS stats against USGS:
```{r}
sitenos <- AppGages$gauge_id
parameterCd <- "00060" 
startDate <- "1981-01-01"  
endDate <- "2010-12-31"

for(i in 1:28)
{
if(i == 1){
  QdatDV <- readNWISdv(sitenos[i], parameterCd, startDate, endDate)
}else{
   QdatDVnew <- readNWISdv(sitenos[i], parameterCd, startDate, endDate) 
  QdatDV <- rbind(QdatDVnew, QdatDV)
   }
}

QdatDV <- QdatDV %>% renameNWISColumns()

QdvStats <- QdatDV %>% 
  group_by(site_no) %>%
  summarize(Mean = mean(Flow),
            Median = median(Flow),
            StandardDeviation = sd(Flow),
            IQR = IQR(Flow),
            TenthP = quantile(Flow, probs = c(0.10)),
            NinetiethP = quantile(Flow, probs = c(0.9)),
            Count = n())

QdvStats <- inner_join(QdvStats, AppGages, by = c("site_no" = "gauge_id")) 

#cfs to mm/day
QdatDV <- QdatDV %>% mutate(Qmm_day = 2.447 * Flow / DRAIN_SQKM)
```

